\documentclass[12pt]{article}

\usepackage[utf8]{inputenc}
\usepackage{geometry}
\geometry{a4paper, margin=1in}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{fancyhdr}
\usepackage{listings} 
\usepackage{xcolor}

\lstdefinelanguage{yaml}{
    keywords={true, false, null, yes, no},
    comment=[l]{\#},
    morestring=[b]',
    morestring=[b]",
    stringstyle=\color{red},
    keywordstyle=\color{blue}\bfseries,
    identifierstyle=\color{black},
    commentstyle=\color{green}\itshape,
    sensitive=false
}

\lstdefinelanguage{vim}{
    keywords={function, endfunction, if, endif, let, set, call, return, elseif},
    keywordstyle=\color{blue}\bfseries,
    sensitive=true,
    morecomment=[l]{\"},   
    morestring=[b]',       
    morestring=[b]"        
}

\lstdefinelanguage{C}{
    keywords={int, float, if, else, while, for, return, void, char, malloc, free, sizeof, struct, NULL},
    keywordstyle=\color{blue}\bfseries,
    sensitive=true,
    morecomment=[l]{//},   
    morecomment=[s]{/*}{*/}, 
    commentstyle=\color{green}\itshape,
    morestring=[b]",       
    morestring=[b]'        
}

\lstdefinelanguage{bash}{
    keywords={if, then, else, elif, fi, for, while, do, done, case, esac, function},
    keywordstyle=\color{blue}\bfseries,
    ndkeywords={echo, exit, read, printf, cd, pwd, mkdir, rm, touch, cp, mv},
    ndkeywordstyle=\color{teal}\bfseries,
    sensitive=true,
    morecomment=[l]{\#},   
    commentstyle=\color{green}\itshape,
    morestring=[b]",       
    morestring=[b]'
}


\lstset{
    basicstyle=\ttfamily\footnotesize, 
    frame=lines, 
    backgroundcolor=\color{white}, 
    rulecolor=\color{black}, 
    tabsize=2, 
    showspaces=false, 
    showstringspaces=false, 
    showtabs=false, 
    breaklines=true, 
    breakatwhitespace=true 
}


\setlength{\headheight}{15pt}
\pagestyle{fancy}
\fancyhf{}
\rhead{Computer Workshop Course}
\lhead{Final Assignment}
\rfoot{Page \thepage}

\title{
    \vspace{2in}
    \textbf{Final Assignment :}\\
    \textbf{Integration of Tools and Practices}\\
    \large Iran University of Science and Technology\\
    \large Department of Computer Engineering\\
    \vspace{2in}
}

\author{
    \vspace{0.5in}
    Qazal Arabali\\
    Computer Workshop\\
    \vspace{0.5in}
}

\date{January 23, 2025}
\begin{document}

\begin{titlepage}
    \maketitle
    \thispagestyle{empty}
\end{titlepage}

\newpage

\tableofcontents
\newpage

\section{Git and GitHub}
    \subsection{Repository Initialization and Commits}
    
To initialize the repository and make the first commits, the following steps were taken :

\begin{enumerate}
    \item Logged into \textbf{GitHub} and created a new repository named :
    \texttt{CW\_Final\_Assignment}
    
    \item Configured the repository settings :
    \begin{itemize}
        \item Set the repository visibility to \textbf{Public}.
        \item Checked the box: \texttt{Initialize this repository with a README}.
    \end{itemize}

    \item Cloned the repository to the local system using the command :
    \begin{lstlisting}
    git clone https://github.com/username/CW_Final_Assignment
    \end{lstlisting}

    \item Navigated to the cloned repository directory :
    \begin{lstlisting}
    cd CW_Final_Assignment
    \end{lstlisting}

    \item Created a new LaTeX file named \texttt{CW\_Final\_Assignment.tex} with the initial structure.

    \item Staged and committed the new file:
    \begin{lstlisting}
    git add CW_Final_Assignment.tex
    git commit -m "Added initial LaTeX file"
    \end{lstlisting}

    \item Pushed the changes to the GitHub repository :
    \begin{lstlisting}
    git push
    \end{lstlisting}
\end{enumerate}

\subsection{GitHub Actions for LaTeX Compilation}

To streamline the process of compiling the LaTeX document and managing different stages of updates, two distinct GitHub Actions workflows were created :

\begin{itemize}
    \item \textbf{Workflow for Intermediate Versions :}
    This workflow compiles the LaTeX document and uploads the resulting PDF as an \textbf{Artifact}. It is used for minor updates or ongoing progress, ensuring the document is accessible for review without cluttering the Releases section.
    
    \item \textbf{Workflow for Major Releases :}
    This workflow compiles the LaTeX document and publishes the resulting PDF as part of a \textbf{GitHub Release}. It is triggered only for significant updates or finalized versions to provide stable releases for public access.
\end{itemize}

\subsubsection*{Workflow for Intermediate Versions (Artifact Only)}

For intermediate updates or minor progress, the LaTeX document is compiled and uploaded as an \textbf{Artifact}. This ensures that ongoing work is accessible for review and testing without cluttering the Releases section.

\begin{enumerate}
    \item \textbf{Trigger :} This workflow runs on every \texttt{push} to the \texttt{main} branch.
    \item \textbf{Outcome :} The compiled PDF is stored in the \textbf{Actions > Artifacts} section for download.
    \item \textbf{Workflow Configuration :}
    \begin{lstlisting}[language=yaml]
        name: Compile LaTeX and Upload Artifact
        
        on:
          push:
            branches:
              - main
        
        jobs:
          build_latex:
            runs-on: ubuntu-latest
            steps:
              - name: Checkout Repository
                uses: actions/checkout@v3
        
              - name: Compile LaTeX Document
                uses: xu-cheng/latex-action@v2
                with:
                  root_file: CW_Final_Assignment.tex
        
              - name: Upload Compiled PDF as Artifact
                uses: actions/upload-artifact@v3
                with:
                  name: CW_Final_Assignment
                  path: ./CW_Final_Assignment.pdf
        \end{lstlisting}
        
\end{enumerate}

\subsubsection*{Workflow for Major Releases}

For significant updates or final versions, the LaTeX document is compiled, and the resulting PDF is published as part of a \textbf{GitHub Release}. This approach ensures that only stable and finalized versions are available in the Releases section.

\begin{enumerate}
    \item \textbf{Trigger :} This workflow runs on every \texttt{push} of a Git tag in the format \texttt{x.x.x} (e.g., \texttt{1.0.0}).
    \item \textbf{Outcome :} The compiled PDF is attached to a new Release in the \textbf{Releases} section.
    \item \textbf{Workflow Configuration :}
    \begin{lstlisting}[language=yaml]
        name: Release Compiled PDF
        
        on:
          push:
            tags:
              - '*.*.*'
        
        jobs:
          build_latex:
            permissions: write-all
            runs-on: ubuntu-latest
            steps:
              - name: Checkout Repository
                uses: actions/checkout@v3
        
              - name: Compile LaTeX Document
                uses: xu-cheng/latex-action@v2
                with:
                  root_file: CW_Final_Assignment.tex
        
              - name: Create Release
                id: create_release
                uses: actions/create-release@v1
                env:
                  GITHUB_TOKEN: ${{ secrets.GITHUB_TOKEN }}
                with:
                  tag_name: ${{ github.ref }}
                  release_name: Release ${{ github.ref }}
                  draft: false
                  prerelease: false
        
              - name: Upload Release Asset
                id: upload_release_asset
                uses: actions/upload-release-asset@v1
                env:
                  GITHUB_TOKEN: ${{ secrets.GITHUB_TOKEN }}
                with:
                  upload_url: ${{ steps.create_release.outputs.upload_url }}
                  asset_path: ./CW_Final_Assignment.pdf
                  asset_name: CW_Final_Assignment.pdf
                  asset_content_type: application/pdf
    \end{lstlisting}
        
\end{enumerate}

\section{Exploration Tasks}
    \subsection{Vim Advanced Features}
    Here are three advanced features of Vim that go beyond the topics covered in class :

    \subsubsection*{Registers}

    Registers in Vim allow you to store multiple pieces of text in separate "clipboard" areas and retrieve them as needed. This can enhance productivity when working with multiple text snippets.

    \begin{itemize}
        \item To copy text to a specific register, use :
        \begin{lstlisting}
        "aY  # Copy text to register 'a'.
        \end{lstlisting}
        \item To paste text from a specific register, use :
        \begin{lstlisting}
        "ap  # Paste text from register 'a'.
        \end{lstlisting}
    \end{itemize}

    \subsubsection*{Macros}

    Macros enable you to record and replay a series of commands, making repetitive tasks easier.

    \begin{itemize}
        \item To start recording a macro :
        \begin{lstlisting}
        q{register}  # Start recording to a register (e.g., 'a').
        \end{lstlisting}
        \item To stop recording :
        \begin{lstlisting}
        q
        \end{lstlisting}
        \item To play back the macro :
        \begin{lstlisting}
        @{register}  # Play the macro stored in the register.
        \end{lstlisting}
    \end{itemize}

    \subsubsection*{Persistent Undo}

    Persistent Undo allows you to keep the undo history of a file even after closing and reopening Vim.

    \begin{itemize}
        \item To enable persistent undo, add the following to your `.vimrc` :
        \begin{lstlisting}[language=vim]
        set undofile
        set undodir=~/.vim/undodir
        \end{lstlisting}
        \item Once enabled, you can navigate the undo history using :
        \begin{lstlisting}
        u       # Undo the last change.
        Ctrl-r  # Redo the last undone change.
        \end{lstlisting}
    \end{itemize}



    \subsection{Memory profiling}
        \subsubsection{Memory Leak}

            A \textbf{memory leak} occurs when a program allocates memory during its execution but fails to release it after it is no longer needed. Over time, this can lead to increased memory usage and potentially cause the program or the system to run out of memory. Memory leaks typically happen due to :

            \begin{itemize}
                \item Forgetting to free dynamically allocated memory.
                \item Retaining references to objects that are no longer needed, preventing their deallocation.
                \item Errors in pointer management, such as overwriting pointers without freeing the associated memory.
            \end{itemize}
            
            For example, in C programming, a memory leak might look like this :
            \begin{lstlisting}[language=c]
            #include <stdlib.h>
            
            int main() {
                int *ptr = (int *)malloc(sizeof(int) * 10); // Memory allocated
                // Forgot to free the allocated memory
                return 0; // Memory is leaked here
            }
            \end{lstlisting}
            
            To prevent memory leaks :
            \begin{itemize}
                \item Always release memory using \texttt{free()} in C or \texttt{delete} in C++.
                \item Use memory profiling tools, such as \textbf{Valgrind}, to identify and fix leaks.
                \item Adopt good coding practices, such as using smart pointers in modern C++.
            \end{itemize}




        \subsubsection{Memory profilers}

            \textbf{Valgrind : Purpose and How It Helps with Memory Leaks}
            
            Valgrind is a powerful instrumentation framework for building dynamic analysis tools. The most commonly used tool within the Valgrind suite is \textbf{Memcheck}, which helps detect memory-related issues in programs, such as memory leaks, use of uninitialized memory, and accessing invalid memory locations.
            
            \paragraph{Purpose of Valgrind}
            \newline
            The primary purpose of Valgrind is to ensure memory correctness in programs by :
            \begin{itemize}
                \item Detecting memory leaks that occur when dynamically allocated memory is not properly freed.
                \item Identifying invalid memory accesses, such as reading or writing outside the bounds of an array.
                \item Detecting the use of uninitialized memory, which can lead to unpredictable behavior.
                \item Helping developers understand and debug complex memory issues.
            \end{itemize}
            
            \paragraph{How Valgrind Helps with Memory Leaks}
            \newline
            Valgrind identifies memory leaks by monitoring all dynamic memory allocations and deallocations. At the end of the program's execution, it reports:
            \begin{itemize}
                \item \textbf{Definite Leaks} : Memory that was allocated but never freed.
                \item \textbf{Indirect Leaks} : Memory that is still reachable through pointers but was not explicitly freed.
                \item \textbf{Suppressed Errors} : Issues that are configured to be ignored.
            \end{itemize}
            
            Using Valgrind, developers can pinpoint the exact location in the code where memory issues occur, making it easier to fix the problem.
            
            \paragraph{Example Usage of Valgrind}
            \newline
            Below is an example of how to use Valgrind to detect memory leaks in a C program :
            
            \begin{lstlisting}[language=C]
            #include <stdlib.h>
            #include <stdio.h>
            
            int main() {
                int *array = (int *)malloc(10 * sizeof(int)); // Allocate memory
                array[0] = 42; // Use the memory
                // Memory is not freed intentionally
                return 0;
            }
            \end{lstlisting}
            \vspace*{2cm}
            To run the program with Valgrind :
            
            \begin{lstlisting}[language=bash]
            valgrind --leak-check=full --track-origins=yes ./my_program
            \end{lstlisting}
            \vspace*{2cm}
            Example output :
            
            \begin{lstlisting}
                ==12345== Memcheck, a memory error detector
                ==12345== LEAK SUMMARY:
                ==12345==    definitely lost: 40 bytes in 1 blocks
                ==12345==    indirectly lost: 0 bytes in 0 blocks
                ==12345==    possibly lost: 0 bytes in 0 blocks
                ==12345==    still reachable: 0 bytes in 0 blocks
                ==12345==         suppressed: 0 bytes in 0 blocks
            \end{lstlisting}
            
            \paragraph{Benefits of Using Valgrind}
            \begin{itemize}
                \item Simplifies debugging of memory-related issues.
                \item Provides detailed error reports for quicker resolution.
                \item Helps improve the stability and performance of software by ensuring proper memory management.
            \end{itemize}
            
            \paragraph{Best Practices When Using Valgrind}
            \begin{itemize}
                \item Compile programs with debugging symbols (\texttt{-g} flag) for detailed reports.
                \item Regularly run Valgrind during development to catch issues early.
                \item Suppress known or irrelevant errors using suppression files.
            \end{itemize}
            

    \subsection{GNU/Linux Bash Scripting}
        \subsubsection{fzf}
            \textbf{What is fuzzy searching ?}

            Fuzzy searching is a technique used to find approximate matches to a given query string, rather than requiring an exact match. This is particularly useful when the user does not remember the full or exact input, or when there are minor typos. Instead of matching exact characters in the correct sequence, fuzzy searching prioritizes likely matches based on the query and displays them in an interactive interface.
            
            \textbf{fzf} (Fuzzy Finder) is a command-line tool that provides an efficient way to perform fuzzy searching. It is highly customizable, interactive, and lightweight. It can be used to search for files, commands, or any list of items in the terminal.
            
            \paragraph{Example usage :}
            \newline
            Suppose you want to search for a file in a directory. With \texttt{fzf}, you simply type a part of the file name, and it will display all matches dynamically as you type.
            
            \begin{lstlisting}[language=bash]
            # Search for a file
            fzf
            
            # Example : Combining with find command
            find . -type f | fzf
            \end{lstlisting}
            
            \paragraph{Key Features of fzf :}
            \begin{itemize}
                \item Interactive interface that displays results as you type.
                \item Supports integration with other tools and commands.
                \item Can be used in combination with pipes for filtering outputs.
                \item Lightweight and customizable with various options.
            \end{itemize}
            
        \subsubsection{Using fzf to find your favorite PDF}
        \subsubsection{Opening the file using Zathura}

\section{Git and FOSS}
    \subsection{README.md}
    \subsection{Issues}
    \subsection{FOSS contribution}

\end{document}
